One-dimensional turbulence (ODT): computationally efficient modeling and simulation of turbulent flows

### 2020-04-27 manuscript/manuscript.tex
Drafted model description section, started software architechture section, outlining impact section. Added triplet map figure.

%
%Questions to answer in this section (from SoftwareX template)
%\begin{enumerate}
%	\item What's the scientific background and motivation for this software?
%	\item Why is this important? What problems does it solve?
%	\item How has the software contributed (and/or how will it contribute in the future) to the process of scientific discovery? Cite papers using the software.
%	\item In what experimental setting might someone use this software?
%	\item What related work is there in the literature?
%	\item What algorithms, other code/software, or ideas are used? Cite them. 
%\end{enumerate}

@@

Describe how ODT works here. Turbulent advection via eddy events implemented by triplet maps. Planar vs. cylindrical configurations. Eddy events happen concurrently with solution of transport equations.  Evolution of transport equations in only one dimension. Limited to one-dimensional flows: jets, wakes, mixing layers. This is fine though because those are common in nature and important to research.

%TEXT COPIED FROM UNFINISHED CONFERENCE PAPER
%
%This study employs the most current version of the ODT code as in Lignell et. al. \cite{Lignell_2018}, and the ODT model in general is described in detail in the literature \cite{Lignell_2018,Lignell_2013,Kerstein_1999,Kerstein_2001,Ashurst_2005}. 
%The ODT code uses a Lagrangian finite-volume formulation for diffusive advancement, including adaptive mesh refinement \cite{Lignell_2013}. In this approach, combustion dilation causes grid cells to increase or decrease in volume while mass remains constant within each cell. 
%Turbulent advection is modeled with stochastic processes called "eddy events" that map functions on the domain via triplet maps. These eddy events occur concurrently with the solution of unsteady one-dimensional transport equations for mass, momentum, and enthalpy. In sooting flame cases, we also transport species mass fractions and soot particle size distribution properties. Because the ODT model is one-dimensional, it is limited to homogeneous or boundary-layer flows, such as jets, wakes, and mixing layers; these types of flows, however, are common in nature and central to turbulence research.
%
% and whose sizes are drawn randomly from a sample distribution. A given eddy of size $l$ and location $x_0$ has an eddy timescale $t$ and an associated eddy rate $1=t$. Eddies occur as a Poisson process in accordance with their given rates. Three ODT parameters control the eddy event process: the eddy rate parameter $C$ scales the rate of occurrence of the eddies; the viscous penalty parameter $Z$ suppresses small eddies; and the large eddy suppression parameter $\beta$ constrains eddies such that they do not reach over the elapsed simulation time.

@@

Things to put in this section (from SoftwareX template)
\begin{itemize}
	\item overview of overall software architecture
	\item optional: pictorial overview
	\item implementation details
	\begin{itemize}
		\item software dependencies?
		\item how to build and run (or leave this in the code README files? maybe brief description here with full details in README.)
		\item how to use post processing tools? (TODO: prune post processing tools as much as possible so there's no excess)
	\end{itemize}
\end{itemize}

Summary of directory structure:
\begin{itemize}
	\item source: source code lives here; yaml files generated by building yaml in build folder get put here as well
		\begin{itemize}
			\item more details on directory structure necessary here? not sure if relevant
		\end{itemize}
	\item build: go here to build the code
		\begin{itemize}
			\item Run: make yaml (only need to do this once, the first time you download/build the code)
			\item set options/paths for cmake in user\_config file
			\item Run: cmake -C user\_config ../source
			\item Run: make -j8
		\end{itemize}
	\item input: input files live here; change input files to specify case details
		\begin{itemize}
			\item each case type has its own folder; make your own if necessary
			\item within each case folder, you should see file input.yaml. this is your all-important input file that makes each case unique
			\begin{itemize}
				\item talk about most important ODT parameters/options here? or leave to documentation?
			\end{itemize}
		\end{itemize}
	\item run: go here to actually run the code; don't forget to make your changes to the input file first
		\begin{itemize}
			\item after building, you'll see odt.x in this folder. you can run it directly (not recommended) or from within one of the bash scripts (better)
			\item don't forget to give your case a name; this name will be used to label data files dumped to the data directory
		\end{itemize}
	\item data: data and runtime files dumped here, organized by case name and individual realizations
		\begin{itemize}
			\item TO DO: raw/processed file types and what they contain, how to use them effectively
		\end{itemize}
	\item post: post processing tools located here, organized by case type
	\item doc: doxygen documentation (if generated) lives here

\end{itemize}

%Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide implementation details.

@@

Present the major functionalities of the software. 
\begin{itemize}
	\item simulating turbulent flow cases, reacting or nonreacting flows
	\item can simulate laminar flow cases too
	\item does these things a whole lot faster than other methods do (LES, DNS especially)
	\item testing LES subgrid modeling assumptions
	\item simulating cases that DNS can't get to because needed simulation length is too long (i.e. late-flame phenomena)
\end{itemize}

Example cases:
\begin{itemize}
	\item From cylindrical ODT paper:
	\begin{enumerate}
		\item pipe flow
		\item nonreacting round jet
		\item round jet flame
	\end{enumerate}
	\item Other possible example cases:
	\begin{itemize}
		\item something to compare with DNS/LES results to illustrate efficiency
		\item no sooting cases since no soot in the code
		\item planar vs. cylindrical comparison
	\end{itemize}
\end{itemize}

Things to talk about with example cases:
\begin{itemize}
	\item compare to DNS/LES cases; computational efficiency, accuracy
	\item how data looks before post processing, how post processing works
\end{itemize}

>>>>

%Questions to answer in this section (from SoftwareX template)
%\begin{enumerate}
%	\item What's the scientific background and motivation for this software?
%	\item Why is this important? What problems does it solve?
%	\item How has the software contributed (and/or how will it contribute in the future) to the process of scientific discovery? Cite papers using the software.
%	\item In what experimental setting might someone use this software?
%	\item What related work is there in the literature?
%	\item What algorithms, other code/software, or ideas are used? Cite them. 
%\end{enumerate}

@@

The ODT model is described in detail in the literature \cite{Kerstein_1999,Kerstein_2001,Ashurst_2005,Lignell_2018,Lignell_2013}; only a brief explanation will be given here. In ODT, turbulent advection is modeled with stochastic processes called eddy events, which punctuate the solution of unsteady, one-dimensional transport equations for mass, momentum, and enthalpy. The ODT code uses a Lagrangian finite-volume formulation for diffusive advancement that includes adaptive mesh refinement \cite{Lignell_2013}. In this approach, mass remains constant inside each grid cell while cell volumes increase or decrease according to dilation. Because the ODT model is one-dimensional, it is limited to homogeneous or boundary-layer flows, such as jets, wakes, and mixing layers; these types of flows, however, are common in nature and central to turbulence research.

Eddy events occur as a Poisson process in accordance with their eddy rates, where a given eddy event of size $l$ and location $x_0$ has an eddy timescale $t$ and an associated eddy rate $1/t$. Three user-defined ODT parameters control the eddy event process: the eddy rate parameter $C$ scales the rate of occurrence of the eddies; the viscous penalty parameter $Z$ suppresses small eddies; and the large eddy suppression parameter $\beta$ constrains eddies such that they do not reach over the elapsed simulation time.

Eddy events modify domain variables using triplet maps, as illustrated for a cylindrical domain in Figure \ref{fig:tripletmap}. For a region of eddy size $l$, the domain is copied to create three map images; the three images are then placed back to back with the middle image inverted to maintain continuity, and the composite is reapplied to the domain. This process applies to all transported variables on the domain. Applied properly, the triplet map increases scalar gradients and decreases length scales consistent with the application of turbulent eddies in real flows, conserves all quantities and their statistical moments, and maintains continuity in property profiles. Subsequent eddies in the same region will result in a cascade of scales, and eddy rates depend on eddy size and the local kinetic energy such that they follow turbulent cascade scaling laws.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../figures/tripletmap/tripletmap.png} 
	\caption{Schematic diagram of a cylindrical triplet map, adapted from \cite{Lignell_2018}. Before the triplet map, the domain contains three grid cells of equal volume, while after the triplet map has been applied, the domain contains nine cells. The nine final cells are labeled according to the cells from which they originated and shaded to indicate that three map images were combined to create the final composite.}
	\label{fig:tripletmap}
\end{figure}

@@

The ODT code is a relatively self-contained C++ package. The system of nonlinear ODEs is solved using CVODE \cite{Hindmarsh_2020} and user input files are processed with YAML \cite{Beder_2008}, both of which are installed locally during the ODT build process. For reacting flow cases, chemical kinetics and transport are handled by Cantera \cite{Goodwin_2018}, which must be previously installed by the user.

%Things to put in this section (from SoftwareX template)
%\begin{itemize}
%	\item overview of overall software architecture
%	\item optional: pictorial overview
%	\item implementation details
%\end{itemize}

@@

%Present the major functionalities of the software. 
%\begin{itemize}
%	\item simulating turbulent flow cases, reacting or nonreacting flows
%	\item can simulate laminar flow cases too
%	\item does these things a whole lot faster than other methods do (LES, DNS especially)
%	\item testing LES subgrid modeling assumptions
%	\item simulating cases that DNS can't get to because needed simulation length is too long (i.e. late-flame phenomena)
%\end{itemize}

%Example cases:
%\begin{itemize}
%	\item From cylindrical ODT paper:
%	\begin{enumerate}
%		\item pipe flow
%		\item nonreacting round jet
%		\item round jet flame
%	\end{enumerate}
%	\item Other possible example cases:
%	\begin{itemize}
%		\item something to compare with DNS/LES results to illustrate efficiency
%		\item no sooting cases since no soot in the code
%		\item planar vs. cylindrical comparison
%	\end{itemize}
%\end{itemize} 
%
%Things to talk about with example cases:
%\begin{itemize}
%	\item compare to DNS/LES cases; computational efficiency, accuracy
%	\item how data looks before post processing, how post processing works
%\end{itemize}
